---
title: "Get started"
author: "Lucas Godoy"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package intend to facilitates the workflow for R users who want to analyze
data from [Portal da Transparência](http://www.portaltransparencia.gov.br/). The
package is based on the section 'download de dados'^[see the [link](http://www.portaltransparencia.gov.br/download-de-dados)].
The [Portal da Transparência](http://www.portaltransparencia.gov.br/) of the 
brazilian federal government has data about the public budget of the country.
Among these data we can analyze the expenses and revenues of the government as well
as data on corporate cards.

## Public budget

In the next piece of code we will visualize the expected
versus realized budget for the science and technology
ministry in 2020.

```{r budget_2020, echo = TRUE, eval = TRUE}
library(transpbrr)
library(data.table)
library(ggplot2)

dt_orc <- download_orcamento(year = 2020)
dt_orc <- dt_orc[nome_orgao_superior == 'Ministerio da Ciencia, Tecnologia, Inovacoes e Comunicacoes',]

dt_orc <- dt_orc[, .(initial = sum(`orcamento_inicial_(r$)`),
                     updated = sum(`orcamento_atualizado_(r$)`),
                     final = sum(`orcamento_realizado_(r$)`)),
                 by = nome_acao]

ggplot(data = dt_orc,
       mapping = aes(x = updated,
                     y = final)) +
    geom_point(alpha = .8) +
    geom_abline(intercept = 0, slope = 1, col = '#357935') +
    theme_bw(base_family = 'AvantGarde') +
    theme(plot.title = element_text(face = 'bold', size = 16),
          panel.grid.major = element_line(linetype = 'longdash'),
          panel.grid.minor = element_line(linetype = 'longdash'),
          strip.background = element_rect(fill = '#357935',
                                          colour = '#357935'),
          strip.text = element_text(color = '#ffffff')) +
    labs(x = 'Budget Expected',
         y = 'Budget Realized',
         title = 'Expected X Realized',
         subtitle = 'Science and Technology Budget')
```

## Public Expenses

### Expenses execution

Loading data on expenses execution for January-2020.

```{r expenses_2020, echo = T, eval = F}
expenses <- download_exec_desp(ym = "202001")
```

### Transfer of resources

Loading data on transfer of resources for January-2014.

```{r transf_2020, echo = T, eval = F}
transfers <- download_transf(ym = "202001")
```

## Corporate Cards

The brazilian government has three kinds of corporate
cards, they are called payment cards (Cartoes de pagamento, in portuguese).
Basically, you can download data from all of these cards using this
function. You just need to specify a year, month and type.
Where type can assume the values:

1. **cpgf**: Data from the federal government payment card
2. **cpcc**: Data from the federal government payment card, used
for centralized purchases
3. **cpdc**: Data from the civil defense payment card

Example:

```{r pc, echo = T, eval = F}
# 1
cpgf <- download_cp(year = 2014, month = 1, type = 'cpgf')
# 2
cpcc <- download_cp(year = 2015, month = 1, type = 'cpcc')
# 3
cpdc <- download_cp(year = 2015, month = 1:2, type = 'cpdc')
```

